---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# indexing

<!-- badges: start -->
<!-- badges: end -->

The `indexing` package provides a convenient function, `indexing()`, to create an index from a variable in a tidy dataset. This index is generated using specified observations as reference points.

## Installation

You can install the development version of indexing from [GitHub](https://github.com/stat545ubc-2023/indexing) with:

``` r
# install.packages("devtools")
devtools::install_github("stat545ubc-2023/indexing")
```

## Demonstrated usage

```{r include=FALSE}
library(dplyr)
library(ggplot2)
library(magrittr)
library(indexing)
library(lubridate)
```

### Example 1:

In this example, the function constructs a time series and establishes an index utilizing the `values` column, with the base observation set at `2023-01-01`.

```{r fig.height=3.5, dpi=300}
# Generate a toy dataset
set.seed(545)
df <- data.frame(date = seq(as.Date("2023-01-01"), 
                            as.Date("2025-12-01"),
                            "month"),
                 values = runif(12*3, 50, 150) * (2 + cumsum(runif(12*3, 0, 0.15))))

# Display the first few rows of the dataset
head(df)

# Generate an index
df2 <- df %>% 
  indexing(values,
           date == as.Date("2023-01-01"))

# Display the first few rows of the indexed dataset
head(df2)

# Compare the original variable vs. the indexed variable
df2 %>% 
  ggplot(aes(x = date)) +
  geom_hline(yintercept = 100,
             linetype = "dashed") +
  geom_line(aes(y = values,
                color = "values")) +
  geom_line(aes(y = index,
                color = "index")) +
  theme_minimal()


```


### Example 2:

Using the toy dataset from earlier, this code chunk generates an indexed variable by averaging values across multiple observations. Specifically, it calculates the average for the year 2023, with a warning issued if this operation was unintended.

```{r fig.height=3.5, dpi=300}
# Generate an index based on the year 2023

df2 <- df %>% 
  indexing(values,
           year(date) == 2023)

# Display the first few rows of the indexed dataset
head(df2)

# Compare the original variable vs. the indexed variable
df2 %>% 
  ggplot(aes(x = date)) +
  geom_hline(yintercept = 100,
             linetype = "dashed") +
  geom_line(aes(y = values,
                color = "values")) +
  geom_line(aes(y = index,
                color = "index")) +
  theme_minimal()

```

### Example 3:

In this example, we demonstrate indexing the GDP per capita for Oceania countries, utilizing the year 1952 as the reference point.

```{r fig.height=3.5, dpi=300}
# Load the gapminder library
library(gapminder)

# Select Oceania countries from the gapminder dataset
oc_gapminder <- gapminder %>% 
  filter(continent == "Oceania") 

# Display the first few rows of the Oceania dataset
oc_gapminder %>% 
  arrange(year) %>% 
  head()

# Create a country-wise index based on GDP per capita for the year 1952
idx_gap <- oc_gapminder %>% 
  group_by(country) %>% 
  indexing(gdpPercap,
           year == 1952) 

# Display the first few rows of the indexed dataset
idx_gap %>% 
  arrange(year) %>% 
  head()

# Compare indexes across countries using a line plot
idx_gap %>% 
  ggplot(aes(x = year, 
             y = index,
             color = country)) +
  geom_line() +
  theme_minimal()

```
